---
title: "A Scalability Analysis of a Multi-agent Framework for solving combinatorial
  optimization via Metaheuristics"
author: Maria Amélia Lopes Silva, Jardell Fillipe da Silva, Sérgio Ricardo de Souza
  e Marcone Jamilson Freitas Souza
date: "UFV - Campus Florestal, CEFET-MG e UFOP 2024"
#output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reshape2)
library(shiny)
library(lmtest)
library(ggplot2)
library(gridExtra)
library(magrittr)
library(dplyr)

rm(list = ls(all = TRUE))

melhor <- read.delim("pmp_instances.txt", header = T)

results <- read.csv("pmp_results.csv", header = T);

tb_mean <- c(mean(results$one_m_g), mean(results$two_m_g), mean(results$four_m_g), mean(results$eight_m_g),mean(results$ten_m_g),mean(results$twenty_m_g),mean(results$thirty_m_g),mean(results$forty_m_g),mean(results$fifty_m_g),mean(results$sixty_m_g)) 
tb_min <- c(mean(results$one_g), mean(results$two_g), mean(results$four_g), mean(results$eight_g),mean(results$ten_g),mean(results$twenty_g),mean(results$thirty_g),mean(results$forty_g),mean(results$fifty_g),mean(results$sixty_g)) 
tb_med <- c(mean(results$one_md_g), mean(results$two_md_g), mean(results$four_md_g), mean(results$eight_md_g), mean(results$ten_md_g), mean(results$twenty_md_g), mean(results$thirty_md_g), mean(results$forty_md_g), mean(results$fifty_md_g), mean(results$sixty_md_g))
tb_mean_time <- c(mean(results$one_m_time), mean(results$two_m_time), mean(results$four_m_time), mean(results$eight_m_time),mean(results$ten_m_time),mean(results$twenty_m_time), mean(results$thirty_m_time), mean(results$forty_m_time), mean(results$fifty_m_time), mean(results$sixty_m_time)) 
tb_med_time <- c(mean(results$one_md_time), mean(results$two_md_time), mean(results$four_md_time), mean(results$eight_md_time),mean(results$ten_md_time),mean(results$twenty_md_time), mean(results$thirty_md_time),mean(results$forty_md_time), mean(results$fifty_md_time), mean(results$sixty_md_time)) 
tb_min_time <- c(mean(results$one_time),mean(results$two_time), mean(results$four_time), mean(results$eight_time), mean(results$ten_time), mean(results$twenty_time), mean(results$thirty_time), mean(results$forty_time), mean(results$fifty_time), mean(results$sixty_time))


```

## **Statistical Analysis UPMSP-ST**

The experiments aim to investigate the performance of the AMAM Framework when the number of agents grows. The objective is to evaluate the structure regarding its scalability. Based on the heterogeneity of the ways of measuring scalability, the experiments assess whether the inclusion of agents in the framework has a direct influence on its performance concerning the quality of the obtained results for the objective function values as well as the obtained runtimes.

### **Gap(%) General**

```{r fitness, echo=FALSE, fig.width=12}
#--------Graph of average gap values (%)--------------------

df <- data.frame(quantidade_agentes = c(1,2,4,8,10,20,30,40,50,60),
                 melhor= tb_min,
                 media = tb_mean,
                 mediana= tb_med )

df <- melt(df,  id.vars = 'quantidade_agentes', variable.name = 'legenda')

plot(df$quantidade_agente,df$value, col = df$legenda, pch = c(rep(19,10),rep(15,10),rep(17,10)), ylab = "gap(%)", xlab = "Number of Agents", xaxt = 'n', title("Behavior of the average gap(%) in relation to the best-known values."));
legend("topright",inset=c(0.02,0.05), legend = c("Best", "Average", "Median"), col=c("black", "red", "green"), pch = c(19,15,17), xpd=TRUE, horiz=TRUE, bty="n")
axis(side = 1, at = c(1, 2, 4, 8, 10, 20, 30, 40, 50, 60), 
     labels = c("One", "Two", "Four", "Eight", "Ten", "Twenty", "Thirty", "Forty", "Fifty", "Sixty"),
     las = 2, cex.axis = 0.8)
```

Behavior of the average gap(%) in relation to the best-known values, the average values, and the medians for the evaluated
scenarios, concerning the number of used agents.

### **General Time**

```{r time, echo=FALSE, fig.width=12}
#--------Graph of average computational time--------------------

time <- results[, c("one_m_time", "two_m_time", "four_m_time", "eight_m_time", "ten_m_time", "twenty_m_time", "thirty_m_time", "forty_m_time", "fifty_m_time", "sixty_m_time")]
average_time <- colMeans(time)
boxplot(time, ylab = "Computational Time", xaxt = 'n');
title("Average Runtime × Number of Agents.")
legend("top",inset=c(0.02,0.0005), legend = c("Average Time"), col=c("red"), pch = c(16), xpd=TRUE, horiz=TRUE, bty="n")
axis(side = 1, at = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), 
     labels = c("One", "Two", "Four", "Eight", "Ten", "Twenty", "Thirty", "Forty", "Fifty", "Sixty"),
     las = 2, cex.axis = 0.8)
points(seq_along(average_time), average_time, col = "red", pch = 16)
```

### **Individual Instances**

```{r ind, echo=FALSE, fig.width=12}

# Interface do Usuário
ui <- fluidPage(
  #titlePanel("Dois Gráficos em um Frame com Shiny"),
  
  sidebarLayout(
    sidebarPanel(
      selectInput("instance", "Instance", 
                  choices = c("I_50_10_S_1-9_1",	"I_50_10_S_1-49_1",	"I_50_10_S_1-99_1",	"I_50_10_S_1-124_1",	"I_50_15_S_1-9_1",	"I_50_15_S_1-49_1",	"I_50_15_S_1-99_1",	"I_50_15_S_1-124_1",	"I_50_20_S_1-9_1",	"I_50_20_S_1-49_1",	"I_50_20_S_1-99_1",	"I_50_20_S_1-124_1",	"I_50_25_S_1-9_1",	"I_50_25_S_1-49_1",	"I_50_25_S_1-99_1",	"I_50_25_S_1-124_1"),
                  selected = "I_50_10_S_1-9_1"),
    ),
    
    mainPanel(
      fluidRow(
        plotOutput("linePlot1", height = "400px"),
        plotOutput("linePlot2", height = "400px")
      )
    )
  )
)

# Lógica do Servidor
server <- function(input, output) {
  output$linePlot1 <- renderPlot({
    
            dados <- read.csv("pmp_data.csv", sep = ",", dec = ".");
            
            d_1 <- dados[dados$instance == input$instance & dados$metodo == "OneAgent", c("fitness", "time")]
            d_2 <- dados[dados$instance == input$instance & dados$metodo == "TwoAgent", c("fitness", "time")]
            d_4 <- dados[dados$instance == input$instance & dados$metodo == "FourAgent", c("fitness", "time")]
            d_8 <- dados[dados$instance == input$instance & dados$metodo == "EightAgent", c("fitness", "time")]
            d_10 <- dados[dados$instance == input$instance & dados$metodo == "TenAgent", c("fitness", "time")]
            d_20 <- dados[dados$instance == input$instance & dados$metodo == "TwentyAgent", c("fitness", "time")]
            d_30 <- dados[dados$instance == input$instance & dados$metodo == "ThirtyAgent", c("fitness", "time")]
            d_40 <- dados[dados$instance == input$instance & dados$metodo == "FortyAgent", c("fitness", "time")]
            d_50 <- dados[dados$instance == input$instance & dados$metodo == "FiftyAgent", c("fitness", "time")]
            d_60 <- dados[dados$instance == input$instance & dados$metodo == "SixtyAgent", c("fitness", "time")]

            melhor <- read.delim("pmp_instances.txt");
            
            dados <-data.frame(one = d_1$fitness,
                   two = d_2$fitness,
                   four = d_4$fitness,
                   eight = d_8$fitness,
                   ten = d_10$fitness,
                   twenty = d_20$fitness,
                   thirty = d_30$fitness,
                   forty = d_40$fitness,
                   fifty = d_50$fitness,
                   sixty = d_60$fitness)

            boxplot(dados, xlab = "Number of Agents", ylab = "Objective Function")

            points.default(1,mean(dados$one),col = "red", pch = 19)
            points.default(1,min(dados$one),col = "blue", pch = 19)
            points.default(1,melhor[melhor$instancia == input$instance, 5],col = "green", pch = 19)
            points.default(2,mean(dados$two),col = "red", pch = 19)
            points.default(2,min(dados$two),col = "blue", pch = 19)
            points.default(2,melhor[melhor$instancia == input$instance, 5],col = "green", pch = 19)
            points.default(3,mean(dados$four),col = "red", pch = 19)
            points.default(3,min(dados$four),col = "blue", pch = 19)
            points.default(3,melhor[melhor$instancia == input$instance, 5],col = "green", pch = 19)
            points.default(4,mean(dados$eight),col = "red", pch = 19)
            points.default(4,min(dados$eight),col = "blue", pch = 19)
            points.default(4,melhor[melhor$instancia == input$instance, 5],col = "green", pch = 19)
            points.default(5,mean(dados$ten),col = "red", pch = 19)
            points.default(5,min(dados$ten),col = "blue", pch = 19)
            points.default(5,melhor[melhor$instancia == input$instance, 5],col = "green", pch = 19)
            points.default(6,mean(dados$twenty),col = "red", pch = 19)
            points.default(6,min(dados$twenty),col = "blue", pch = 19)
            points.default(6,melhor[melhor$instancia == input$instance, 5],col = "green", pch = 19)
            points.default(7,mean(dados$thirty),col = "red", pch = 19)
            points.default(7,min(dados$thirty),col = "blue", pch = 19)
            points.default(7,melhor[melhor$instancia == input$instance, 5],col = "green", pch = 19)
            points.default(8,mean(dados$forty),col = "red", pch = 19)
            points.default(8,min(dados$forty),col = "blue", pch = 19)
            points.default(8,melhor[melhor$instancia == input$instance, 5],col = "green", pch = 19)
            points.default(9,mean(dados$fifty),col = "red", pch = 19)
            points.default(9,min(dados$fifty),col = "blue", pch = 19)
            points.default(9,melhor[melhor$instancia == input$instance, 5],col = "green", pch = 19)
            points.default(10,mean(dados$sixty),col = "red", pch = 19)
            points.default(10,min(dados$sixty),col = "blue", pch = 19)
            points.default(10,melhor[melhor$instancia == input$instance, 5],col = "green", pch = 19)
            legend("topright",inset=c(0,0.018), legend = c("Agent Average", "Best of Agent", "Best of Literature"), col=c("red", "blue", "green"), pch = 19, xpd=TRUE, horiz=TRUE, bty="n")
            })
  
  output$linePlot2 <- renderPlot({
    
    dados <- read.csv("pmp_data.csv", sep = ",", dec = ".");
    
    d_1 <- dados[dados$instance == input$instance & dados$metodo == "OneAgent", c("fitness", "time")]
    d_2 <- dados[dados$instance == input$instance & dados$metodo == "TwoAgent", c("fitness", "time")]
    d_4 <- dados[dados$instance == input$instance & dados$metodo == "FourAgent", c("fitness", "time")]
    d_8 <- dados[dados$instance == input$instance & dados$metodo == "EightAgent", c("fitness", "time")]
    d_10 <- dados[dados$instance == input$instance & dados$metodo == "TenAgent", c("fitness", "time")]
    d_20 <- dados[dados$instance == input$instance & dados$metodo == "TwentyAgent", c("fitness", "time")]
    d_30 <- dados[dados$instance == input$instance & dados$metodo == "ThirtyAgent", c("fitness", "time")]
    d_40 <- dados[dados$instance == input$instance & dados$metodo == "FortyAgent", c("fitness", "time")]
    d_50 <- dados[dados$instance == input$instance & dados$metodo == "FiftyAgent", c("fitness", "time")]
    d_60 <- dados[dados$instance == input$instance & dados$metodo == "SixtyAgent", c("fitness", "time")]
    
    tempo <-data.frame(one = d_1$time,
                   two = d_2$time,
                   four = d_4$time,
                   eight = d_8$time,
                   ten = d_10$time,
                   twenty = d_20$time,
                   thirty = d_30$time,
                   forty = d_40$time,
                   fifty = d_50$time,
                   sixty = d_60$time)

    boxplot(tempo, ylab = "Time(s)", xlab = "Number of Agents")
  })
}

# Rodar o Aplicativo Shiny
shinyApp(ui = ui, server = server)
```

The graphs demonstrate the behavior of the AMAM Framework in relation to the objective function and execution time as the number of agents increases for each tested instance.

### **Two Axles**

```{r twoaxles, echo=FALSE, fig.width=12}

#-------------------Two Axles ---------------------------

rm(list = ls(all = TRUE))

dados_time <- read.csv("pmp_data_time.csv", dec = ".");
dados_g <- read.csv("pmp_data_fitness.csv", dec = ".");

agentes <- c(1, 2, 4, 8, 10, 20, 30, 40, 50, 60)

fo <- c(mean(dados_g$one), mean(dados_g$two), mean(dados_g$four), mean(dados_g$eight), mean(dados_g$ten), mean(dados_g$twenty), mean(dados_g$thirty), mean(dados_g$forty), mean(dados_g$fifty), mean(dados_g$sixty))
tempo <- c(mean(dados_time$one), mean(dados_time$two), mean(dados_time$four), mean(dados_time$eight), mean(dados_time$ten), mean(dados_time$twenty), mean(dados_time$thirty), mean(dados_time$forty), mean(dados_time$fifty), mean(dados_time$sixty))

## add extra space to right margin of plot within frame
par(mar=c(4, 5, 4, 5) + 0.1)

## Plot first set of data and draw its axis
plot(agentes, fo, pch=16, axes=FALSE, #ylim=c(-14,-2), 
     xlab="", ylab="", 
     type="b",col="black")
axis(2, col="black",las=1)
mtext("Objective Function (average)",side=2.5,line=3.5)

## Allow a second plot on the same graph
par(new=TRUE)

## Plot the second plot and put axis scale on right
plot(agentes, tempo, pch=15,  xlab="", ylab="", 
     axes=FALSE, type="b", col="red")
mtext("Average Time(s)",side=4,col="red",line=3)
axis(4, col="red",col.axis="red",las=1)
nomes_ingles <- c("one", "two", "four", "eight", "ten", "twenty", "thirty", "forty", "fifty", "sixty")
axis(side=1, at=agentes, labels=nomes_ingles, las = 2, cex.axis = 0.8)
mtext("Number of Agents",side=1,col="black",line=2.5) 

legend("topleft",inset=c(0.4,-0.01),legend=c("Objective Function","Time"), text.col=c("black","red"),pch=c(16,15),col=c("black","red"),  xpd=TRUE, horiz=TRUE, bty="n")

```

The graph compare the average values of the objective function and runtime for the ten evaluated scenarios using
a graph with two ordinate axes.

### **Statistical Analysis**

First, we checked if parametric tests could be applied to the collected data by evaluating homoscedasticity, independence, and normality. Based on the results, it was determined that parametric tests could not be applied to either the objective function values or the computational time value.

#### Possibility of applying a parametric test - Objective Function

```{r sa, echo=FALSE, fig.width=12}
#------------------Parametric-----------------

rm(list = ls(all = TRUE))

dados <- read.csv("pmp_data.csv", sep = ",", dec = ".");

#Premises for carrying out the parametric time test

model <- aov(fitness ~ metodo, data = dados);

#Durbin-Watson to check independence
dwtest(model)

#Fligner-Killeen to check homoscedasticity
fligner.test (fitness ~ metodo, data = dados)

#Shapiro-Wilk to check normality
#shapiro.test(resid(model))
qqnorm(resid(model))
qqline(resid(model))
```

#### Possibility of applying a parametric test - Computational Time

```{r sa_time, echo=FALSE, fig.width=12}
#Premises for carrying out the parametric time test

modelt <- aov(time ~ metodo, data = dados);

#Durbin-Watson to check independence
dwtest(modelt)

#Fligner-Killeen to check homoscedasticity
fligner.test (time ~ metodo, data = dados)

#Shapiro-Wilk to check normality
#shapiro.test(resid(model))
qqnorm(resid(modelt))
qqline(resid(modelt))
```

Since the used metaheuristics are algorithms with a stochastic nature, each of the evaluated scenarios in these experiments was executed 30 times for each instance. The results obtained in each of these scenarios were compared using the Kruskal-Wallis non-parametric hypothesis test, with a confidence level of 95%. After verifying whether there was a statistical difference in each comparison, the Wilcoxon test was used to compare pairs of populations.

#### Non-parametric statistical test Objective Function

```{r naosa, echo=FALSE, fig.width=12}
#------------------Non-Parametric-------------

rm(list = ls(all = TRUE))

dados <- read.csv("pmp_data.csv", sep = ",", dec = ".")

met <- c("OneAgent", "TwoAgent", "FourAgent", "EightAgent", "TenAgent", "TwentyAgent", "ThirtyAgent", "FortyAgent", "FiftyAgent", "SixtyAgent")
dados$metodo <- factor(dados$metodo, levels = met)
#dados[order(dados$metodo), ]

#show("Teste estatístico Função Objetivo")
kruskal.test(dados$fitness ~ dados$metodo)
pairwise.wilcox.test(dados$fitness, dados$metodo, paired=T, p.adjust.method="bonferroni",alternative="less")
pairwise.wilcox.test(dados$fitness, dados$metodo, paired=T, p.adjust.method="bonferroni",alternative="greater")
```

#### Non-parametric statistical test  Computational Time

```{r naosatime, echo=FALSE, fig.width=12}
#show("Teste estatístico Tempo Computacional")
kruskal.test(dados$time ~ dados$metodo)
pairwise.wilcox.test(dados$time, dados$metodo, paired=T, p.adjust.method="bonferroni", alternative="less")
pairwise.wilcox.test(dados$time, dados$metodo, paired=T, p.adjust.method="bonferroni", alternative="greater")
```

### **Linear Adjustment**

A linear adjustment analysis of the objective function and runtime values was carried out concerning the inclusion of agents. Thus, a logarithmic scale was adopted for both the evaluated values (objective function and runtime) and the number of agents to verify whether the data would assume linear behavior in this situation.

```{r ld, echo=FALSE, fig.width=12}

rm(list = ls(all = TRUE))

results <- read.csv("pmp_results.csv", header = T);

tb_mean <- c(mean(results$one_m_g), mean(results$two_m_g), mean(results$four_m_g), mean(results$eight_m_g),mean(results$ten_m_g),mean(results$twenty_m_g),mean(results$thirty_m_g),mean(results$forty_m_g),mean(results$fifty_m_g),mean(results$sixty_m_g)) 
tb_min <- c(mean(results$one_g), mean(results$two_g), mean(results$four_g), mean(results$eight_g),mean(results$ten_g),mean(results$twenty_g),mean(results$thirty_g),mean(results$forty_g),mean(results$fifty_g),mean(results$sixty_g)) 
tb_med <- c(mean(results$one_md_g), mean(results$two_md_g), mean(results$four_md_g), mean(results$eight_md_g), mean(results$ten_md_g), mean(results$twenty_md_g), mean(results$thirty_md_g), mean(results$forty_md_g), mean(results$fifty_md_g), mean(results$sixty_md_g))
tb_mean_time <- c(mean(results$one_m_time), mean(results$two_m_time), mean(results$four_m_time), mean(results$eight_m_time),mean(results$ten_m_time),mean(results$twenty_m_time), mean(results$thirty_m_time), mean(results$forty_m_time), mean(results$fifty_m_time), mean(results$sixty_m_time)) 
tb_med_time <- c(mean(results$one_md_time), mean(results$two_md_time), mean(results$four_md_time), mean(results$eight_md_time),mean(results$ten_md_time),mean(results$twenty_md_time), mean(results$thirty_md_time),mean(results$forty_md_time), mean(results$fifty_md_time), mean(results$sixty_md_time)) 
tb_min_time <- c(mean(results$one_time),mean(results$two_time), mean(results$four_time), mean(results$eight_time), mean(results$ten_time), mean(results$twenty_time), mean(results$thirty_time), mean(results$forty_time), mean(results$fifty_time), mean(results$sixty_time))

df <- data.frame(quantidade_agentes = c(1,2,4,8,10,20,30,40,50,60),
                 melhor= tb_min,
                 media = tb_mean,
                 mediana= tb_med )

df <- melt(df,  id.vars = 'quantidade_agentes', variable.name = 'legenda')

#--------Graph of average gap values (%) - logarithmic scale--------------------

plot(log(df$quantidade_agentes),log(df$value), col = df$legenda, pch = c(rep(19,10),rep(15,10),rep(17,10)), ylab = "Log gap(%)", xlab = "Log Number of Agents", xaxt = 'n');
legend("topright",inset=c(0.02, 0.01), legend = c("Best", "Average", "Median"), col=c("black", "red", "green"), pch = c(19,15,17), xpd=TRUE, horiz=TRUE, bty="n")
axis(side=1, at = round(log(1),2),las=2)
axis(side=1, at = round(log(2),2),las=2)
axis(side=1, at = round(log(4),2),las=2)
axis(side=1, at = round(log(8),2),las=2)
axis(side=1, at = round(log(10),2),las=2)
axis(side=1, at = round(log(20),2),las=2)
axis(side=1, at = round(log(30),2),las=2)
axis(side=1, at = round(log(40),2),las=2)
axis(side=1, at = round(log(50),2),las=2)
axis(side=1, at = round(log(60),2),las=2)

#--------Linear Regression--------------------

df <- data.frame(quantidade_agentes = c(log(1),log(2),log(4),log(8),log(10),log(20),log(30),log(40),log(50),log(60)),
                 melhor= log(tb_min),
                 media = log(tb_mean),
                 mediana= log(tb_med),
                 melhor_time = tb_min_time,
                 media_time = tb_mean_time,
                 mediana_time= tb_med_time)

modelo_media <- lm(data = df, formula = media ~ quantidade_agentes)
#modelo_mediana <- lm(data = df, formula = mediana ~ quantidade_agentes)
#modelo_melhor <- lm(data = df, formula = melhor ~ quantidade_agentes)
modelo_media_time <- lm(data = df, formula = media_time ~ quantidade_agentes)
modelo_mediana_time <- lm(data = df, formula = mediana_time ~ quantidade_agentes)
modelo_melhor_time <- lm(data = df, formula = melhor_time ~ quantidade_agentes)
```

#### Linear Regression gap values (%)

```{r ldfo, echo=FALSE, fig.width=12}
#--------Linear Regression gap values (%)--------------------

#--------LR Average----------

modelo_media$coefficients
summary(modelo_media)

#--------LR Median----------

#modelo_mediana$coefficients
#summary(modelo_mediana)

#--------LR Best----------

#modelo_melhor$coefficients
#summary(modelo_melhor)

g1 <- ggplot(df, aes(x=quantidade_agentes, y=media)) + geom_point() + ggtitle("Average") + ylab("") + xlab("") + geom_smooth(method = lm, se = FALSE)
#g2 <- ggplot(df, aes(x=quantidade_agentes, y=mediana)) + geom_point() + ggtitle("") + ylab("") + xlab("") + geom_smooth(method = lm, se = FALSE)
#g3 <- ggplot(df, aes(x=quantidade_agentes, y=melhor)) + geom_point() + ggtitle("") + ylab("") + xlab("") + geom_smooth(method = lm, se = FALSE)

#grid.arrange(g1, g2, g3, nrow = 3)

g1
```

#### Linear Regression Computational Time

```{r ldtime, echo=FALSE, fig.width=12}
#--------Linear Regression Computational Time--------------------

#--------LR Average----------

modelo_media_time$coefficients
summary(modelo_media_time)

#--------LR Median----------

modelo_mediana_time$coefficients
summary(modelo_mediana_time)


#--------LR Best----------

modelo_melhor_time$coefficients
summary(modelo_melhor_time)

t1 <- ggplot(df, aes(x=quantidade_agentes, y=media_time)) + geom_point() + ggtitle("Average") + ylab("") + xlab("") + geom_smooth(method = lm, se = FALSE)
t2 <- ggplot(df, aes(x=quantidade_agentes, y=mediana_time)) + geom_point() + ggtitle("Median") + ylab("") + xlab("") + geom_smooth(method = lm, se = FALSE)
t3 <- ggplot(df, aes(x=quantidade_agentes, y=melhor_time)) + geom_point() + ggtitle("Best") + ylab("") + xlab("") + geom_smooth(method = lm, se = FALSE)

grid.arrange(t1, t2, t3, nrow = 3)

```

### **Single agent x Multi-agent**

This section presents the tests carried out to compare the performances of the \texttt{AMAM} Framework with a single agent and with sixty agents. This comparison involves evaluating the values of the objective function and runtime in four different situations:

1. the objective function values obtained with 1800 executions of the single-agent system $\times$ the objective function values obtained by each of the 60 agents of each of the 30 executions of the multi-agent system (1800 $\times$ 1800);

2. the objective function values obtained by 30 executions of the single-agent system $\times$ the objective function values obtained with 30 executions of the multi-agent system; in this case, the best solution found in each execution of the multi-agent system with 60 agents was used (30 $\times$ 30);

3. the objective function values obtained with 1800 executions of the single-agent system $\times$ the objective function values obtained with 30 executions of the multi-agent system, in this case being used the best solution found in each execution of the multi-agent system with 60 agents (1800 $\times$ 30);

4. the objective function values obtained from the execution of 60 single agents, selecting the best objective function value (best of the machine) and worst computational time $\times$ value obtained with 30 executions of the multi-agent system (best of 60 agents) ( 30 $\times$ 60).

```{r multi, echo=FALSE, fig.width=12}

melhor <- read.delim("pmp_instances.txt", header = T)

#------------1800 x 1800 ------------------

dados_1800_one <- read.csv("pmp_50_1_60.csv", sep = ",", dec = ".");
dados_1800_sixty <- read.csv("pmp_50_1_60_60.csv", sep = ",", dec = ".");
dados_1800_sixty <- dados_1800_sixty %>%
  mutate(metodo = ifelse(metodo == 60, "SixtyAgent", metodo))

dados_1800_one <- merge(dados_1800_one, melhor, by.x = "instance", by.y = "instance")
dados_1800_one$gap <- (dados_1800_one$fitness - dados_1800_one$best) / dados_1800_one$best

dados_1800_sixty <- merge(dados_1800_sixty, melhor, by.x = "instance", by.y = "instance")
dados_1800_sixty$gap <- (dados_1800_sixty$fitness - dados_1800_sixty$best) / dados_1800_sixty$best

d_1_1800 <- data.frame(instancia = character(),
                       tratamento = character(),
                       time = double(),
                       results = double(),
                       gap = double());

d_1_1800 <- data.frame(instancia = dados_1800_one$instance,
                       tratamento = dados_1800_one$metodo,
                       time = dados_1800_one$time,
                       results = dados_1800_one$fitness,
                       gap = dados_1800_one$gap);

d_60_1800 <- data.frame(instancia = character(),
                        tratamento = character(),
                        time = double(),
                        results = double(),
                        gap = double());

d_60_1800 <- data.frame(instancia = dados_1800_sixty$instance,
                        tratamento = dados_1800_sixty$metodo,
                        time = dados_1800_sixty$time,
                        results = dados_1800_sixty$fitness,
                        gap = dados_1800_sixty$gap);

d_60_1800 <- rbind.data.frame(d_1_1800, d_60_1800);

#--------------30 x 30-------------

dados_30 <- read.csv("pmp_data.csv", sep = ",", dec = ".");
dados_30 <- merge(dados_30, melhor, by.x = "instance", by.y = "instance")
dados_30$gap <- (dados_30$fitness - dados_30$best) / dados_30$best

d_1_30 <- data.frame(instancia = character(),
                     tratamento = character(),
                     time = double(),
                     results = double(),
                     gap = double());

d_1_30 <- data.frame(instancia = subset(dados_30, metodo %in% c("OneAgent", "SixtyAgent"), select = instance),
                     tratamento = subset(dados_30, metodo %in% c("OneAgent", "SixtyAgent"), select = metodo),
                     time = subset(dados_30, metodo %in% c("OneAgent", "SixtyAgent"), select = time),
                     results = subset(dados_30, metodo %in% c("OneAgent", "SixtyAgent"), select = fitness),
                     gap = subset(dados_30, metodo %in% c("OneAgent", "SixtyAgent"), select = gap));

colnames(d_1_30) <- c("instancia", "tratamento", "time", "results", "gap")

#--------------1800 x 30-------------

d_1800_30 <- rbind.data.frame(subset(d_1_30, tratamento == "SixtyAgent"), subset(d_1_1800, tratamento == "OneAgent"));

#-------------30 x 60-----------------

d_1_30_m <- data.frame(instancia = character(),
                       tratamento = character(),
                       time = double(),
                       results = double(),
                       gap = double());

d_1_30_m <- d_1_1800 %>%
  group_by(instancia) %>%
  mutate(grupo = rep(1:1800, each = 60, length.out = n())) %>%  # Criando um grupo de 60 observações para cada instância
  group_by(instancia, grupo) %>%
  summarise(results = min(results),
            time = max(time),
            gap = gap[which.min(results)],
            tratamento = tratamento[which.min(results)]) %>%
  filter(grupo != 31) %>%  # Removendo as linhas onde grupo é igual a 31
  ungroup() %>%
  select(-grupo)

d_1_30_m <- rbind.data.frame(d_1_30_m, subset(d_1_30, tratamento == "SixtyAgent"));

p = ggplot(d_60_1800, aes(tratamento, gap)) + geom_boxplot() + #ggtitle("1800 x 1800 - Objective Function") +   
  labs(x = "number of agents", y = "gap(%)")
t = ggplot(d_60_1800, aes(tratamento, time)) + geom_boxplot() + #ggtitle("1800 x 1800 - Computational Time") +   
  labs(x = "number of agents", y = "time(s)")

p1 = ggplot(d_1_30, aes(tratamento, gap)) + geom_boxplot() + #ggtitle("30 x 30 - Objective Function") +   
  labs(x = "number of agents", y = "gap(%)")
t1 = ggplot(d_1_30, aes(tratamento, time)) + geom_boxplot() + #ggtitle("30 x 30 - Computational Time") +   
  labs(x = "number of agents", y = "time(s)")

p2 = ggplot(d_1800_30, aes(tratamento, gap)) + geom_boxplot() + #ggtitle("1800 x 30 - Objective Function") +   
  labs(x = "number of agents", y = "gap(%)")
t2 = ggplot(d_1800_30, aes(tratamento, time)) + geom_boxplot() + #ggtitle("1800 x 30 - Computational Time") +   
  labs(x = "number of agents", y = "time(s)")

p3 = ggplot(d_1_30_m, aes(tratamento, gap)) + geom_boxplot() + #ggtitle("Machine 30 x 30 - Objective Function") +   
  labs(x = "number of agents", y = "gap(%)")
t3 = ggplot(d_1_30_m, aes(tratamento, time)) + geom_boxplot() + #ggtitle("Machine 30 x 30 - Computational Time") +   
  labs(x = "number of agents", y = "time(s)")
```

#### Situation 1 - 1800 x 1800

```{r multi1, echo=FALSE, fig.width=12}
#_________________________________________________________TESTE ESTATISTICO_________________________________________________

#show("Situation 1 - 1800 x 1800")
grid.arrange(p, t, nrow = 1)

pairwise.t.test(d_60_1800$gap, d_60_1800$tratamento, p.adjust.method="bonferroni")
pairwise.t.test(d_60_1800$time, d_60_1800$tratamento, p.adjust.method="bonferroni")
```

#### Situation 2 - 30 x 30

```{r multi2, echo=FALSE, fig.width=12}
#show("Situation 2 - 30 x 30")
grid.arrange(p2, t2, nrow = 1)

pairwise.t.test(d_1800_30$gap, d_1800_30$tratamento, p.adjust.method="bonferroni")
pairwise.t.test(d_1800_30$time, d_1800_30$tratamento, p.adjust.method="bonferroni")
```

#### Situation 3 - 1800 x 30

```{r multi3, echo=FALSE, fig.width=12}
#show("Situation 3 - 1800 x 30")
grid.arrange(p1, t1, nrow = 1)

pairwise.t.test(d_1_30$gap, d_1_30$tratamento, paired=T, p.adjust.method="bonferroni")
pairwise.t.test(d_1_30$time, d_1_30$tratamento, paired=T, p.adjust.method="bonferroni")
```

#### Situation 4 - 30 x 60

```{r multi4, echo=FALSE, fig.width=12}
#show("Situation 4 - 30 x 60")
grid.arrange(p3, t3, nrow = 1)

pairwise.t.test(d_1_30_m$gap, d_1_30_m$tratamento, p.adjust.method="bonferroni")
pairwise.t.test(d_1_30_m$time, d_1_30_m$tratamento, p.adjust.method="bonferroni")
```